Time-variable surface mapping of exoplanets: an application to Io

### 2020-06-16 paper/paper.tex
Added some figures, sarted writing the data section.

% !TeX root = ./ms.tex
\documentclass[modern]{aastex62}

@@

%
\section{Introduction}
The surface of Jupiter's moon Io is littered with hundreds of volcanoes appearing as bright spots in the near infrared and varying in intensity on timescales of days to decades.
The volcanic activity is driven by tidal interactions with Jupiter and sustained by the Laplace resonance with Europa and Ganymede \citep{peale_melting_1979}.
Io is a perfect analogue of an extremely volcanically active exoplanet.
Such rocky exoplanets, often dubbed \emph{super-Ios} or \emph{lava worlds}, have gathered a lot of interest in recent years because the photometric and spectral signatures of volcanic activity on such worlds will likely be detectable in the near future \citep{kaltenegger_detecting_2010,henning_highly_2018,oza_sodium_2019}.
Studying volcanism on both Io and exoplanets is scientifically valuable because it provides a window into the properties of early Earth and it informs theories of planet formation.

@@

\section{Data}
\label{sec:data}
\section{Model}
\label{sec:model}
% Discuss starry, pixels vs. schmixels vs. harmonics, forward model for static map, NMF model for dynamic map
\subsection{Orbital parameters}
\label{ssec:orbital_parameters}
% Talk about JPL horizons, uncertainties on parameters, geometry of occultations etc.

\section{The inverse problem}
\label{sec:inverse_problem}
% bulk of the results in the paper
\subsection{The information content of a light curve}
% information in different kinds of occultations/phase curves, reflected light vs. emitted light, posterior shrinkage etc.

\label{ssec:information_content}
\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=0.5\linewidth]{figures/information_content.pdf}
    \oscaption{InformationContent}{%
        This is a plot of a pretty function. And at the end of this
        caption is a symbol with a link to the \emph{exact} script
        that generated it, hosted on \textsf{GitHub}.
        \label{fig:information_content}
    }
    \end{centering}
\end{figure}

@@

>>>>

% !TeX root = ./paper.tex
\documentclass[modern]{aastex62}

@@

%
\section{Introduction}
%% ADD PARAGRAPH REVIEWING PREVIOUS LITERATURE ON MAPPING EXOPLANETS, MENTION JWST AND LUVOIR
%% MAYBE EVEN IN THE ABSTRACT
The surface of Jupiter's moon Io is littered with hundreds of volcanoes appearing as bright spots in the near infrared and varying in intensity on timescales of days to decades.
The volcanic activity is driven by tidal interactions with Jupiter and sustained by the Laplace resonance with Europa and Ganymede \citep{peale_melting_1979}.
Io is a perfect analogue of an extremely volcanically active exoplanet.
Such rocky exoplanets, often dubbed \emph{super-Ios} or \emph{lava worlds}, have gathered a lot of interest in recent years because the photometric and spectral signatures of volcanic activity on such worlds will likely be detectable in the near future \citep{kaltenegger_detecting_2010,henning_highly_2018,oza_sodium_2019}.
Studying volcanism on both Io and exoplanets is scientifically valuable because it provides a window into the properties of early Earth and it informs theories of planet formation.

@@

\section{Data}
\label{sec:data}

\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=0.5\linewidth]{figures/irtf_max_flux.pdf}
    \oscaption{irtf_dataset}{%
        Maximum flux for each curve taken using NASA's IRTF telescope during an
        occultation of Io by Jupiter.
       \label{fig:irtf_max_flux}
    }
    \end{centering}
\end{figure}

The dataset we use to infer maps of Io's surface consists of 112 observations of Io in the near infrared taken from 1996 until 2018 using various instruments at NASA's IRTF observatory.
Figure~\ref{fig:irtf_max_flux} shows the maximum flux in $\mathrm{GW}/\mu \mathrm{m}/\mathrm{sr}$ for each light curve plotted as a function of time which approximately corresponds to the disc integrated flux of Io at the beginning of ingress or the end of egress prior to it being occulted by Jupiter.
Two features of this plot are apparent.
First although the observations span decades the cadence is non-uniform with only a few observations taken between 2008 and 2016.
Second, the baseline brightness is varying stochastically by a large amount and there are several notable events of increased volcanic activity.
All of the observations were taken while Io was \emph{in eclipse}, meaning that it was in Jupiter's shadow and all of the observed emission from the surface is due to thermal radiation.
Observations of Io \emph{in sunlight} on the other hand probe both the thermal (volcanic) emission and the surface albedo variations in the near infrared.

\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=0.5\linewidth]{figures/irtf_sample_lightcurves.pdf}
    \oscaption{irtf_dataset}{%
        A selection of sample light curves taken during occultations of Io by Jupiter in our dataset.
        The step-like morphology of the light curves is due to bright volcanoes on Io's surface coming in or out of view during an occultation.
        These light curves thus visibly encode information about the features on the surface.
        \label{fig:irtf_sample_lightcurves}
    }
    \end{centering}
\end{figure}

In Figure~\ref{fig:irtf_sample_lightcurves} we display a random subset of all the light curves
in the dataset.
All occultations last for $\sim4$ min and the cadence for each light varies but it on the order of a second.
It follows that over the course of a single exposure Jupiter's limb move by about 15km which provides a lower bound for the size of the features on the surface that we can reliably estimate.
The shapes of light curves strongly deviate from the smooth variability one would expect assuming a homogeneous distribution of thermal emission on the surface.
Especially prominent are light curves with clear step-like features which are present when bright spots come in or out of view during the course of an occultation.
The fact that these features are so clearly visible means that even individual light curves encode a wealth of information about the surface.

The photometric quality varies from year to year because multiple instruments were used over the years.
An additional issue is that estimated errorbars aren't provided.
As with all ground based photometry, the observations are influenced by atmospheric variability which induces correlated noise in the light curves.
Because of this the flux isn't always monotonically increasing or decreasing as one would expect.
All of these issues need to be accounted for in the final model.

\section{Model}
\label{sec:model}
% Discuss starry, pixels vs. schmixels vs. harmonics, forward model for static map, NMF model for dynamic map
\subsection{Orbital parameters}
\label{ssec:orbital_parameters}
To model the occultations, first we need to know the trajectory of Jupiter relative to Io for any given time.
To compute the trajectory we use  \href{https://ssd.jpl.nasa.gov/horizons.cgi}{JPL Horizons database} and the latest planetary ephemeris DE430 \citep{folkner_planetary_2014}.
The quantities given by the Horizons database needed to uniquely define the relative trajectory
of Jupiter and the orientation of Io with respect to the observer on Earth are given in Table~\ref{tab:horizons}.
JPL Horizons interface provides $(\alpha,\delta)$ coordinates of both objects and the
central longitude and latitude

% Talk about JPL horizons, uncertainties on parameters, geometry of occultations etc.

\section{The inverse problem}
\label{sec:inverse_problem}
% bulk of the results in the paper
\subsection{The information content of a light curve}
\label{ssec:information_content}
% information in different kinds of occultations/phase curves, reflected light vs. emitted light, posterior shrinkage etc.

\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=0.5\linewidth]{figures/information_content.pdf}
    \oscaption{InformationContent}{%
        This is a plot of a pretty function. And at the end of this
        caption is a symbol with a link to the \emph{exact} script
        that generated it, hosted on \textsf{GitHub}.
        \label{fig:information_content}
    }
    \end{centering}
\end{figure}

@@

\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=1.\linewidth]{figures/single_event_fit.pdf}
    \oscaption{single_event_fit}{%
        This is a plot of a pretty function. And at the end of this
        caption is a symbol with a link to the \emph{exact} script
        that generated it, hosted on \textsf{GitHub}.
        \label{fig:single_event_fit}
    }
    \end{centering}
\end{figure}
